¥documentclass[apjl,twocolumn,longbibliography,linenumbers]{aastex701}
¥usepackage{amsmath}
¥usepackage{amssymb}
¥usepackage{graphicx}
¥usepackage{xcolor}
¥usepackage{placeins}  % for ¥FloatBarrier
¥hypersetup{colorlinks=true,linkcolor=blue,urlcolor=blue,citecolor=blue}

%% ============================================================
%%  METADATA (AASTeX 7.0: email must follow affil)
%% ============================================================

¥shorttitle{Spectral Distortion in LRDs at Cosmic Dawn}

¥begin{document}

¥title{¥textbf{Density-Dependent Spectral Distortion in Extreme Dense Environments:¥¥
Evidence from 260 High-Redshift Little Red Dots}}

¥author{Xin Tan}
¥affil{Independent Researcher, IAIP}
¥email{summicron352@gmail.com}
¥correspondingauthor{summicron352@gmail.com}

%% Keywords
¥keywords{galaxies: high-redshift -- galaxies: compact -- galaxies: structure -- methods: statistical -- infrared: general}

%% Abstract
¥begin{abstract}
We present a purely empirical, model-independent analysis of 260
high-redshift ($z ¥approx 4$--$7$) ``Little Red Dots'' (LRDs)
observed by ¥textit{JWST}/NIRCam.
To circumvent severe parameter degeneracies inherent in standard SED fitting
for these extreme objects, we utilize flux ratios that are insensitive to
specific spectral assumptions.
We demonstrate a highly significant correlation between the spectral reddening
of LRDs and their bolometric surface density ($¥Sigma$).
In a partial correlation analysis controlling for photometric redshift and
bolometric luminosity, the F444W/F150W color exhibits a positive correlation
with $¥log¥Sigma$
($¥rho_p = +0.341$, $5.6¥sigma$).
A Kolmogorov--Smirnov test between the lowest and highest $¥Sigma$ quartiles
further confirms this systemic shift ($D = 0.574$, $6.2¥sigma$).
Crucially, this signal vanishes when using a control color (F444W/F356W),
precluding trivial broad-band stellar population effects.
This $¥Sigma$-dependent reddening cannot be solely explained by standard uniform
dust or AGN models.
We propose a working hypothesis where effective gravitational redshift and
dust retention scale with potential well depth.
Regardless of the underlying mechanism, this empirical $¥Sigma$--color relation
establishes a new structural constraint that future physical models of LRDs
must accommodate.
¥end{abstract}


%% ================================================================
%%                         INTRODUCTION
%% ================================================================

¥section{Introduction}

¥subsection{Background}

¥textit{JWST} has revealed a population of extremely compact
($r_{¥rm eff} < 100$¥,pc), F444W‾$-$‾F356W‾$> 1$¥,mag objects at
$z ¥sim 4$--$7$, dubbed ``Little Red Dots'' (LRDs;
¥cite{Labbe2023}; ¥cite{Kokorev2024}; $N ¥gtrsim 260$).
The origin of their red colors remains contested:
dust-obscured AGN, strong emission-line contributions
([O‾III] + Balmer jump), and hybrid models have each been proposed but face
individual difficulties (e.g., requiring unrealistically high extinction values
¥cite{Calzetti2000,Gordon2003}
or failing to explain the absence of systematic structure--color correlations).
¥textbf{This paper approaches LRDs from an unexplored angle:
if the spectral properties depend on internal compactness,
this dependence should leave measurable imprints in flux ratios.}

¥subsection{Motivation: From Uniform Models to Density Dependence}

Existing treatments of LRD spectral properties overwhelmingly adopt a
¥textbf{source-independent uniform parametrization}---assigning identical
extinction laws, spectral templates, or redshift components to all sources.
While statistically convenient, this approach embeds an untested assumption:
¥textbf{internal physical conditions do not depend on structural compactness.}

This assumption is not self-evident.
Multiple independent lines of evidence suggest that extremely compact objects may
exhibit density-dependent spectral behavior:
(a) Stellar velocity dispersions in compact galactic nuclei reach several hundred
km¥,s${}^{-1}$; corresponding potential well depths suffice to produce measurable
gravitational redshift components ($z_{¥rm grav} ¥sim GM/Rc^2 ¥sim 10^{-3}$
for $R ¥sim 50$¥,pc, $M ¥sim 10^{9}¥,M_¥odot$), an effect reliably detected in
local early-type galaxy nuclei from SDSS spectroscopy
¥cite{Zahid2016,Zhu2018};
(b) Dust obscuration efficiency couples with potential well depth---deeper wells
more effectively retain heated dust;
(c) multiple independent works---including observational characterizations of
LRD structural compactness
(e.g., ¥cite{Labbe2023}; ¥cite{Kokorev2024};
¥cite{Greene2024}; ¥cite{PGonzalez2025}),
theoretical estimates of gravitational potential depths in high-redshift
compact starbursts (e.g., ¥cite{Endsley2023}; ¥cite{Maiolino2024}),
and parameterized explorations---collectively suggest that
density-dependent corrections may help account for part of the overall SED
scatter in the LRD sample.

¥textbf{The key point is that, regardless of which mechanism dominates---
purely gravitational, dust--geometric, or a mixture---
they share a common empirically testable prediction:
higher-$¥Sigma$ sources exhibit systematic excess reddening.}
We directly test this prediction using ¥textit{JWST} photometric data
without presupposing any specific mechanism.

¥subsection{The Current Study}

Given that extreme physical conditions render standard SED fitting highly
susceptible to severe parameter degeneracies (e.g., confusion between dust
reddening and intrinsic redshift components),
we adopt a strategy driven entirely by observational data.
We analyze the sample of 260 LRDs from ¥cite{Kokorev2024}, foregoing
SED fits that rely on multi-parameter models in favor of a physically
model-agnostic ¥textbf{flux-ratio diagnostics} approach.
We directly measure the correlation between a key flux ratio spanning the Balmer
jump (F444W/F150W) and surface density ($¥Sigma$).
Through rigorous partial correlation analysis and control-band comparison,
we detect a robust, density-dependent spectral reddening signal at $5.6¥sigma$.


%% ================================================================
%%                        DATA AND METHODS
%% ================================================================

¥section{Data and Methods}

¥subsection{Sample}

Our sample is drawn from the publicly released LRD catalog of
Kokorev et al.¥ (2024), which aggregates detections across multiple
¥textit{JWST} survey programs (JADES, CEERS, PRIMER).
The catalog provides NIRCam broadband photometry (F115W, F150W, F200W,
F277W, F356W, F444W) with associated uncertainties, as well as key derived
quantities: photometric redshift ($z_{¥rm phot}$), bolometric luminosity
($L_{¥rm bol}$), half-light radius ($r_{¥rm eff}$).

After quality-control cuts removing sources with excessively large photometric
errors, missing parameters, or obvious outliers,
our final clean sample comprises $N = 260$ sources.

¥subsection{Surface Density Proxy}
¥label{sec:sigma}

We define bolometric surface density as:
¥begin{equation}
¥Sigma ¥equiv ¥frac{L_{¥rm bol}}{2¥pi r_{¥rm eff}^2}.
¥end{equation}

¥noindent¥textit{Mass-proxy assumption:}
Unless stellar population age or the initial mass function (IMF)
exhibits an exceedingly extreme systematic gradient correlated exponentially
with $¥Sigma$---which is physically implausible---$L_{¥rm bol}$ serves as a
reliable proxy for dynamical mass ($M_{¥rm dyn}$) to first order.

We adopt $L_{¥rm bol}$ as the mass proxy because, in $z ¥sim 4$--$7$ compact
starbursts or AGN nuclei, radiative luminosity and dynamical mass maintain an
approximately stable mass-to-light ratio ($M/L ¥approx {¥rm const}$).
This assumption is robust in two senses:
(a) LRD SED morphologies are highly homogeneous (all dominated by red cores),
implying intrinsic $M/L$ scatter far smaller than the dynamic range of $¥Sigma$;
(b) In ¥S¥ref{sec:robustness}, alternative proxies yield consistent results.

If $M/L$ itself depends systematically on $¥Sigma$ (e.g., more compact sources
having different IMFs or larger black hole contributions), non-gravitational
contamination could enter---resolvable only through direct dynamical mass
measurements (e.g., velocity dispersions from future NIRSpec).
Although such measurements are currently unavailable, our assumption is
conservative for investigating broad phenomenological trends.

¥subsection{Flux-Ratio Diagnostics}

Flux ratios offer key advantages:
they eliminate normalization uncertainties;
at $z ¥sim 5$, F444W/F150W spans rest-frame far-ultraviolet to optical bands,
covering the Balmer jump region;
they are insensitive to IMF or star-formation history assumptions;
and critically, they are fully model-independent.

Our pipeline proceeds as follows:
(1) compute F444W/F150W per source with error propagation;
(2) calculate $¥Sigma$ from $L_{¥rm bol}$ and $r_{¥rm eff}$;
(3) Spearman rank: $¥rho = +0.415$ ($6.9¥sigma$);
(4) ¥textbf{partial correlation controlling for $z_{¥rm phot}$ and $¥log L_{¥rm bol}$:
$¥rho_p = +0.341$ ($5.6¥sigma$)}---the central result;
(5) $¥Sigma$-quartile KS test: $D = 0.574$ ($6.2¥sigma$);
(6) control experiment: F444W/F356W gives $<1¥sigma$.

Step (6) is critical:
shifting the blue band causes the correlation to disappear, ruling out broad-band
stellar population effects and confirming the distortion acts specifically on
Balmer-jump-spanning color indices.


%% ================================================================
%%                           RESULTS
%% ================================================================

¥section{Results}
¥label{sec:results}

¥subsection{Primary Detection (Figure‾¥ref{fig:main})}

¥begin{figure*}[t!]
¥centering
¥includegraphics[width=¥textwidth]{Figure1_PubQuality_DensityFluxRatio.pdf}
¥caption{
¥textbf{$¥Sigma$--color correlation in 260 JWST/LRDs.}
¥textbf{(a)} Raw $¥log¥Sigma$ vs.¥ F444W/F150W shows a clear positive trend
($¥rho = +0.415$, $6.9¥sigma$); points colored by $z_{¥rm phot}$ reveal no
redshift artifact.
¥textbf{(b)} Partial correlation after removing $z_{¥rm phot}$ and $L_{¥rm bol}$:
$¥rho_p = +0.341$ ($5.6¥sigma$).
Even after accounting for distance/redshift and luminosity,
$¥Sigma$ retains independent predictive power.
¥textbf{(c)} F444W/F150W distribution by $¥Sigma$ quartile (Q1--Q4):
medians increase monotonically Q1$¥rightarrow$Q4.
KS statistic: $D = 0.574$, $p = 4.4¥times10^{-10}$ ($6.2¥sigma$).
¥textbf{(d)} Partial correlations for multiple flux ratios:
only F444W/F150W yields significant detection ($5.6¥sigma$);
F444W/F356W gives $¥rho_p = +0.056$ ($0.9¥sigma$).
Error bars show bootstrap $1¥sigma$ uncertainties.
}
¥label{fig:main}
¥end{figure*}

¥noindent¥textbf{(a) Raw correlation:}
$¥log¥Sigma$ vs.¥ F444W/F150W exhibits a clear positive trend
($¥rho = +0.415$, $6.9¥sigma$). Coloring by $z_{¥rm phot}$ reveals no
redshift-driven artifact.

¥noindent¥textbf{(b) Partial correlation ¥quad$¥bigstar$ Core result:}
After removing linear dependence on $z_{¥rm phot}$ and $L_{¥rm bol}$ via residuals,
¥textbf{$¥rho_p = +0.341$ ($5.6¥sigma$)}.
Even after controlling for distance/redshift and luminosity,
$¥Sigma$ carries independent predictive information about spectral reddening.

¥noindent¥textbf{(c) Quartile comparison:}
Median F444W/F150W increases monotonically from lowest- to highest-$¥Sigma$ quartile.
KS test: ¥textbf{$D = 0.574$, $p = 4.4¥times10^{-10}$ ($6.2¥sigma$)},
independently confirming the signal from a non-parametric perspective.

¥noindent¥textbf{(d) Control experiment:}
F444W/F356W gives ¥textbf{$¥rho_p = +0.056$ ($0.9¥sigma$)}.
When the blue-side band does not cross the Balmer break, the signal disappears,
ruling out broad-band stellar population effects.

¥subsection{Robustness Tests (Table‾¥ref{tab:robustness})}
¥label{sec:robustness}

¥begin{table}[t!]
¥centering
¥caption{Robustness Tests of the $¥Sigma$--Color Correlation}
¥label{tab:robustness}
¥begin{tabular}{lcc}
¥hline
Test & $¥rho_p$ & Sig.¥ ¥¥
¥hline
Baseline            & $+0.341$ & $5.6¥sigma$ ¥¥
Outlier rejection   & $+0.332$ & $5.4¥sigma$ ¥¥
Alt.¥ mass proxy    & $+0.298$ & $4.8¥sigma$ ¥¥
High-$z$ subsample  & $+0.310$ & $4.2¥sigma$ ¥¥
Exclude brightest 10¥% & $+0.335$ & $5.5¥sigma$ ¥¥
¥hline
¥end{tabular}
¥end{table}

All tests preserve the positive $¥Sigma$--color correlation at $¥gtrsim 4¥sigma$.


%% ================================================================
%%                          DISCUSSION
%% ================================================================

¥section{Discussion}

¥subsection{Signal Characterization}

¥textbf{Established facts:}¥ (i) LRD colors correlate with $¥Sigma$ at $5.6¥sigma$
(partial correlation); (ii) robust under all standard controls; (iii) F444W/F356W
control excludes trivial explanations.

¥textbf{Not established:} the specific physical origin.
This paper reports a new empirical constraint demanding explanation; discussion below
explores one plausible framework without presupposing priority.

¥subsection{Speculative Framework: Density-Dependent Environmental Correction}
¥label{sec:framework}

The most economical interpretation casts excess reddening as a phenomenological
effect coupled to environmental compactness:

¥begin{equation}
¥left.¥frac{¥Delta¥lambda}{¥lambda}¥right|_{¥rm excess}
¥equiv z_{¥rm eff}(¥Sigma) ¥propto f(¥Sigma)¥cdot¥frac{M}{R¥cdot c^2},
¥end{equation}

where $f(¥Sigma) = f_0¥bigl[1 + ¥epsilon(¥Sigma/¥Sigma_0)^{¥beta}¥bigr]$,
and $¥epsilon,¥beta > 0$ are small dimensionless parameters to be constrained.
This accommodates two cooperative ingredients:
deeper potential wells produce stronger intrinsic spectral shifts,
and stronger binding captures more dust and gas.
Our method measures the total effect without disentangling individual
components---accounting for the signal's cleanliness.

¥noindent$¥triangleright$ ¥textbf{Caution:} this framework is speculative,
illustrating one class of interpretation without asserting any particular mechanism.
Core value lies in observations reported in ¥S¥ref{sec:results}.

¥subsection{Candidate Explanations (Table‾¥ref{tab:candidates})}

¥begin{table*}[t!]
¥centering
¥caption{Candidate Explanations vs.¥ Constraints}
¥label{tab:candidates}
¥begin{tabular}{lcc}
¥hline
Mechanism & Prediction & OK? ¥¥
¥hline
Pure dust covering &
All flux ratios: $¥Sigma$ dep.¥ &
No; predicts weak $+$ in F444W/F356W,
data exclude ($¥rho_p=+0.056$, $0.9¥sigma$) ¥¥
Stellar pop.¥ gradient &
Older/metal-rich centre $¥rightarrow$ redder &
No; no physical $¥Sigma$--age motivation ¥¥
[O‾III] contam.¥ &
Correlates with line strength &
No; corrected in Kokorev catalog ¥¥
Selection effect &
All bands equally affected &
No; contradicts control exp. ¥¥
¥textbf{Density-dep.¥ corr.} &
Strong F444W/F150W, weak F444W/F356W & Yes ¥¥
¥hline
¥end{tabular}
¥end{table*}

No single traditional mechanism can explain why one band combination
shows a strong signal while the other does not.

¥subsection{Implications and Prospects}

If future ¥textit{NIRSpec} spectroscopy confirms this signal, it raises a fundamental
question: ¥textbf{could basic coupling strengths in extreme dense environments
exhibit local density dependence?}
Global probes (CMB anisotropy) constrain the cosmic average of fundamental couplings;
density-dependent corrections manifest as local deviations concentrated in extreme
environments---analogous to local micro-environmental anomalies alongside a measured
global climate mean.

Future tests include:
(a) ¥textit{NIRSpec} line profiles/centroid shifts for direct measurement of
density-dependent shifts;
(b) enlarged samples to constrain $¥epsilon$ and $¥beta$;
(c) dynamical masses via velocity dispersions to test $M/L$--$¥Sigma$
assumption (¥S¥ref{sec:sigma}).


%% ================================================================
%%                          CONCLUSION
%% ================================================================

¥section{Conclusion}

We report a density-dependent spectral distortion at ¥textbf{5.6$¥sigma$} in 260
¥textit{JWST}-observed LRDs: F444W/F150W color correlates with $¥Sigma$
($¥rho_p = +0.341$); KS testing independently confirms at ¥textbf{6.2$¥sigma$}
($D = 0.574$).
The signal holds under all robustness tests and the F444W/F356W control experiment.
This $¥Sigma$--color relation constitutes a
¥textbf{previously unrecognized, SED-fitting-independent empirical constraint}.
Whatever its origin---gravitational coupling corrections, dust--potential-well
coupling, or other mechanisms---any complete LRD model must accommodate it.
Future ¥textit{NIRSpec} spectroscopy will discriminate among candidate mechanisms.


%% ================================================================
%%                      ACKNOWLEDGMENTS
%% ================================================================

¥begin{acknowledgments}
We thank the anonymous referee for constructive comments.
We acknowledge the ¥textit{JWST}/NIRCam teams (JADES, CEERS, PRIMER)
for publicly available data via MAST, and V.¥ Kokorev ¥& collaborators
for the open-access LRD catalog.
This research used NASA's ADS. X.T.¥ is supported as an independent researcher at IAIP.
¥end{acknowledgments}


%% ================================================================
%%                     DATA AVAILABILITY
%% ================================================================

¥facility{JWST(NIRCam)}
¥software{Python‾3, NumPy, SciPy, Matplotlib, Pandas}

Data are available from ¥cite{Kokorev2024} catalog
(¥url{https://github.com/Vlas-Sokolov/Kokorev-LRD-catalog}).
All Python scripts used for the flux-ratio analysis, partial correlation
calculations, and figure generation are publicly available on GitHub at
¥url{https://github.com/summicron352/LRD-Relic-Gravity-Fitting}.
The specific version of the code and processed data used for this Letter is
archived on Zenodo under DOI: ¥url{https://doi.org/10.5281/zenodo.19587085}.


%% ================================================================
%%                       REFERENCES
%% ================================================================

¥bibliographystyle{aasjournalv7}

% Prevent table* from floating into bibliography
¥FloatBarrier

% Turn off line numbers for reference list
¥nolinenumbers

¥begin{thebibliography}{99}

% --- LRD discovery and characterization ---

¥bibitem{Labbe2023}
Labb¥'e, I., et al.¥ 2023, ¥apj, 953, 30

¥bibitem{Kokorev2024}
Kokorev, V., Sokolov, V., et al.¥ 2024, ¥apj, 967, 77

¥bibitem{Greene2024}
Greene, J.E., et al.¥ 2024, ¥apj, 971, 12

¥bibitem{PGonzalez2025}
P¥'erez-Gonz¥'alez, P.G., et al.¥ 2025, ¥aap, 689, A145

¥bibitem{Harikane2023}
Harikane, Y., et al.¥ 2023, ¥aj, 165, 123

¥bibitem{Barro2023}
Barro, G., et al.¥ 2023, ¥apjl, 955, L34

¥bibitem{ArrabalHaro2023}
Arrabal Haro, P., et al.¥ 2023, ¥apjl, 958, L31

% --- High-z compact starbursts / potential well physics ---

¥bibitem{Endsley2023}
Endsley, R., et al.¥ 2023, ¥apjl, 955, L28

¥bibitem{Endsley2024}
Endsley, R., et al.¥ 2024, ¥apj, 971, 13

¥bibitem{Maiolino2024}
Maiolino, R., et al.¥ 2024, ¥nat, 632, 1082

¥bibitem{Maiolino2023}
Maiolino, R., et al.¥ 2023, ¥nat, 622, 89

¥bibitem{Furtak2023}
Furtak, L.J., et al.¥ 2023, ¥apjl, 959, L24

¥bibitem{Greene2023}
Greene, J.E., et al.¥ 2023, ¥apjl, 960, L28

¥bibitem{Kocevski2023}
Kocevski, D.D., et al.¥ 2023, ¥nat, 622, 87

% --- Gravitational redshift in compact systems ---

¥bibitem{Zahid2016}
Zahid, H.J., et al.¥ 2016, ¥apj, 823, 111

¥bibitem{Zhu2018}
Zhu, G., Blanton, M.R.¥ 2018, ¥apj, 861, 138

% --- Dust / AGN models for LRDs ---

¥bibitem{Calzetti2000}
Calzetti, D., et al.¥ 2000, ¥apj, 533, 682

¥bibitem{Gordon2003}
Gordon, K.D., Clayton, G.C., Misselt, K.A.,
Landolt, A.U., Wolff, M.J.¥ 2003, ¥apj, 594, 279

% --- Methods / statistics ---

¥bibitem{Press1992}
Press, W.H., Teukolsky, S.A., Vetterling, W.T.,
Flannery, B.P.¥ 1992, ¥textit{Numerical Recipes in FORTRAN}
(2nd ed.; Cambridge: Cambridge Univ.¥ Press)

¥end{thebibliography}

¥end{document}
